---
title: "Ofav-SH-NH4_withCopy"
author: "AnaPalacio"
date: "May 4th, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


# General project set-up 

```{r libraries , results="hide"}

    getwd()

    # Get all libraries and sources required to run the script
        source("STEPoneFunction.R")# R.Cunning steponeR function
        library(dplyr)
        library(plyr)
        library(reshape2)
        library(ggplot2)
        library(nlme)
        library(lme4)
        #library(lmerTest)
```

# 1. COPY NUMBER CALCULATION FOR THE HOST 

Four qPCR plates were run with sybr green using HOST-actin and single copy primers. A fith plate was done after re-extraction of samples that appeared to be outliers. CTs were analized with R.Cunning steponeR function to get the Ofav Actin copy number:

* Get list of plate files to read
* Calculate the copy number ratios. 

 
```{r CalculateCopyNumber}

Ofav.copyList <- list.files(path="HostCopyNumber/data", pattern="csv", full.names=T)
  # Ofav.copy


  # Run stepone function to get Ratios
  
  OfavCopy.Out <- steponeR (files=Ofav.copyList, target.ratios=c("Of_Actin.Of_SC"), 
                       fluor.norm=list(Of_Actin=0, Of_SC=0),
                       copy.number=list(Of_Actin=1, Of_SC=1),
                       ploidy=list(Of_Actin=2, Of_SC=2),
                       extract=list(Of_Actin=0.982, Of_SC=0.982))
  
  # Target ratio results
  OfavCopy<-OfavCopy.Out$result

```
### A. Data CLEANING

```{r RemoveNTC}
    # 1. Check and remove NTC wells 
      ntc <- OfavCopy[which(OfavCopy$Sample.Name=="0-NTC"), ]
      OfavCopy <- droplevels(OfavCopy[!rownames(OfavCopy) %in% rownames(ntc), ])
      
    # 2. remove samples with 1 amplificaation 
      NoData <- OfavCopy[which((OfavCopy$Of_Actin.reps==1) |  (OfavCopy$Of_SC.reps==1)), ]
      OfavCopy <- droplevels(OfavCopy[!rownames(OfavCopy) %in% rownames(NoData), ])
      
    # 3. or high SD
      StDe2 <- OfavCopy[which((OfavCopy$Of_Actin.CT.sd)>0.6|(OfavCopy$Of_SC.CT.sd>0.6)), ]
      OfavCopy <- droplevels(OfavCopy[!rownames(OfavCopy) %in% rownames(StDe2), ])
      
    # Remove columns with # of reps
      OfavCopy <- subset( OfavCopy, select = -c(Of_Actin.reps, Of_SC.reps) )
      
    # 4. Chose bw samples ran more than once. Some samples looked like outliers. Runned 2x and also were run from 2nd extractions (CopyNUmber PLate5)
      # * First, I create a unique ID for each sample.run and count the number
      # of times that each core was ran in the qPCR. Then I export the data from
      # the samples that have duplicates in a .csv file that I "manually" check
    OfavCopy$ID<-paste(OfavCopy$Sample.Name,OfavCopy$File.Name, sep = ".")
    
  # Check the number of times that a sigle sample was ran  
    ReRunA_Copy <- OfavCopy[duplicated(OfavCopy$Sample.Name),]
    
    n_RunA_Copy <- data.frame(table(OfavCopy$Sample.Name))
    colnames(n_RunA_Copy)<-c("Sample.Name","RanA")
    OfavCopy<-join(OfavCopy, n_RunA_Copy, type = "left")
    
    DuplicatesA_Copy <- OfavCopy[(OfavCopy$RanA>1),]
    write.csv(DuplicatesA_Copy, file = 'DuplicatesA_Copy.csv')
    
```

* Look at the duplicates in 'DuplicatesA_Copy.csv' and select the best runs per core. In the DuplicatesA file I delete the good runs and keep the bad ones that are saved a new file called "ToRemove_Copy.csv". I used this file to remove them from the general data frame.

* Finally check for duplicates again (you should not have) and summarize the data to check that it makes sense.

```{r}
  # Remove bad replicates
    ToRem_Copy<-read.csv("ToRemove_Copy.csv") 
    # LAst checked 5/04/17
    OfavCopy<-OfavCopy[!(OfavCopy$ID %in% ToRem_Copy$ID),]
    
  # Check for replicates again--should have none
    
    n_RunB_Copy <- data.frame(table(OfavCopy$Sample.Name))
    ReRunB_Copy <- OfavCopy[duplicated(OfavCopy$Sample.Name),]
    colnames(n_RunB_Copy)<-c("Sample.Name","RanB")
    OfavCopy<-join(OfavCopy, n_RunB_Copy, type = "left")
    
    
  # List of dupplicated samples, should have 0 rows now
    DuplicatesB_Copy <- OfavCopy[(OfavCopy$RanB>1),]
    # write.csv(DuplicatesB, file = 'DuplicatesC.csv')
```
### B. Calculate mean copy number by colony

Get Colony number from the sample.name

```{r ColonyLabels}
    colony.core <- rbind.fill(lapply(strsplit(as.character(OfavCopy$Sample.Name), split="-"), 
        function(X) data.frame(t(X))))
    colnames(colony.core) <- c("Colony", "Core")
    OfavCopy <- cbind(colony.core, OfavCopy)
```
Graph mean +- SD copy number and CTS and SD of technical replicates 

```{r meanCNumberAndCTs}

  CopyNumber <- ggplot(OfavCopy, aes (Colony, Of_Actin.Of_SC, colour=factor(Colony)))
  CopyNumber + stat_summary(fun.data = "mean_cl_boot",geom = "errorbar", width = 0.2)+
    stat_summary(fun.y=mean, geom="point", size =3, alpha=0.5) +
    geom_jitter()
  
  par(mfrow=c(2,2))
      # Mean CTs and sd bu Colony
      plot(x=OfavCopy$Colony, y=OfavCopy$Of_Actin.CT.mean)
      plot(x=OfavCopy$Colony, y=OfavCopy$Of_SC.CT.mean)
      plot(x=OfavCopy$Colony, y=OfavCopy$Of_Actin.CT.sd)
      plot(x=OfavCopy$Colony, y=OfavCopy$Of_SC.CT.sd)
  par(mfrow=c(1,1))
```


```{r MeanCopyPerColony}
ddply(OfavCopy, c("Colony"), summarise,
      mean = mean(Of_Actin.Of_SC), sd = sd(Of_Actin.Of_SC), se =sd(Of_Actin.Of_SC)/sqrt(length(Of_Actin.Of_SC)))
```
  
    
```{r ANOVAcoloniesCopy}
  CopyANOVA <-aov (Of_Actin.Of_SC~Colony, data=OfavCopy)
  summary(CopyANOVA)

  TukeyHSD(CopyANOVA, ordered = TRUE)
```

Since there is not differences bw Colonies 4 and 6, and 7 and 8, pool all the data from the pair of colonies and calculate the mean copy number for them? Keep 4 and 6 with different values?

```{r MeanCopies}
 Copy1 <- OfavCopy[which((OfavCopy$Colony=="4")|(OfavCopy$Colony=="6")), ]
   # summary(Copy1$Of_Actin.Of_SC)
   mean(Copy1$Of_Actin.Of_SC)
   sd(Copy1$Of_Actin.Of_SC)

 Copy2 <- OfavCopy[which((OfavCopy$Colony=="7")|(OfavCopy$Colony=="8")), ]
   # summary(Copy2)
    mean(Copy2$Of_Actin.Of_SC)
    sd(Copy2$Of_Actin.Of_SC)
   
 
```
Use 2 for colonies 4, 3 for and 6, and 8 for colonies 7 and 8. SD for colonies 7 and 8 is very high, rerun samples with min and max values?



##---------------------------------------------------------------------------------
# 2. Calculate qPCR RATIOS (Symbiont/Coral) 

Get the raw data for Ofav R.Cunning steponeR function:

* Get list of plate files to read

* Calculate the ratios

* Extract the results
 
```{r CalculateRatios}

  Ofav.plates <- list.files(path="data", pattern="csv", full.names=T)
  # Ofav.plates

  # Run stepone function to get Ratios
  
  Ofav.Out <- steponeR(files=Ofav.plates, target.ratios=c("A.Ofav", "B.Ofav","C.Ofav", "D.Ofav"), 
                       fluor.norm=list(A=-0.064, B=4.197, C=3.798, D=0, Ofav=7.416),
                       copy.number=list(A=1, B=1,C=50, D=3, Ofav=1),
                       ploidy=list(A=1, B=1, C=1, D=1, Ofav=2),
                       extract=list(A=0.813, B=0.813, C=0.813, D=0.813, Ofav=0.982))
  
  # Target ratio results
  Ofav<-Ofav.Out$result

```

### Data CLEANING part A

```{r RemoveNTCandControls}
    # 1. Check and remove NTC wells
    
    ntc <- Ofav[which(Ofav$Sample.Name=="NTC"), ]
    Ofav <- droplevels(Ofav[!rownames(Ofav) %in% rownames(ntc), ])
    
    # 2. Check and remove + Control wells
    
    Positive <- Ofav[which(Ofav$Sample.Name=="Control"), ]
    Ofav <- droplevels(Ofav[!rownames(Ofav) %in% rownames(Positive), ])
```


### Calculate total S/H ratio, clade specific ratios and clade (D) proportion

```{r}

# 1. If Clade only detected in one technical replicate, set its ratio to NA (becomes zero)
    Ofav$A.Ofav[which(Ofav$A.reps==1)] <- NA
    Ofav$B.Ofav[which(Ofav$B.reps==1)] <- NA
    Ofav$C.Ofav[which(Ofav$C.reps==1)] <- NA
    Ofav$D.Ofav[which(Ofav$D.reps==1)] <- NA

# 2. Rename cols and make NA=0
    colnames(Ofav)[which(colnames(Ofav) %in% c("A.Ofav", "B.Ofav","C.Ofav", "D.Ofav" ))] <- c("A.SH", "B.SH","C.SH", "D.SH")  

    Ofav$A.SH[is.na(Ofav$A.SH)] <- 0
    Ofav$B.SH[is.na(Ofav$B.SH)] <- 0
    Ofav$C.SH[is.na(Ofav$C.SH)] <- 0
    Ofav$D.SH[is.na(Ofav$D.SH)] <- 0
    
# 3. Get the ratios, correct host copy number and calculate log 10(ratios)
    # Total ratio
    Ofav$tot.SH <- Ofav$A.SH + Ofav$B.SH + Ofav$C.SH + Ofav$D.SH  
    
    #Host copy number correction
    
    colony.core2 <- rbind.fill(lapply(strsplit(as.character(Ofav$Sample.Name), split="-"), 
        function(X) data.frame(t(X))))
    colnames(colony.core2) <- c("Colony", "Core")
    Ofav <- cbind(colony.core2, Ofav)
    
    Ofav$HostCopy <-NA
    Ofav$HostCopy[which(Ofav$Colony==4)] <- 2
    Ofav$HostCopy[which(Ofav$Colony==6)] <- 3
    Ofav$HostCopy[which(Ofav$Colony==7)] <- 8
    Ofav$HostCopy[which(Ofav$Colony==8)] <- 8
    
    Ofav$tot.SH_c <- (Ofav$tot.SH/(1/Ofav$HostCopy))
    
    Ofav$Log_tot.SH <- log10(Ofav$tot.SH)  # Calculate log10 SH ratio
    Ofav$Log_tot.SH[which(Ofav$tot.SH==0)] <- NA
    
    Ofav$Log_tot.SH_c <- log10(Ofav$tot.SH_c)  # Calculate log10 SH ratio
    Ofav$Log_tot.SH_c[which(Ofav$tot.SH_c==0)] <- NA
  
# 4. Clade Proportion
      # D Proportion
      Ofav$D.Prp<-(Ofav$D.SH/Ofav$tot.SH)
      D.Pro<-hist(Ofav$D.Prp)
```

All the cores are cleary dominated by D symbionts, so I won't perform any analysis splitting the cores by clade and will use the total SH cell ratio, assuming this is basically D1. 


### DATA CLEANING part B

3. Remove unused colomns and check and remove samples that did not have calculated SH cell ratio

```{r CleaningB}

    Ofav <- subset( Ofav, select = -c
                    (Core, A.CT.mean, B.CT.mean, A.CT.sd, B.CT.sd, A.reps, B.reps, A.SH, B.SH) )

    NoHSratio <- Ofav[which(Ofav$tot.SH_c==0), ]
    Ofav <- droplevels(Ofav[!rownames(Ofav) %in% rownames(NoHSratio), ])
```    

4. Chose bw samples ran more than once. This is a manual step in which I compare the quality of the amplifications and decide which run to keep depending on the SD between technical replicates and CTs. 
 
* First, I create a unique ID for each sample.run and count the number of times that each core was ran in the qPCR. Then I export the data from the samples that have duplicates in a .csv file that I "manually" check

```{r}
    Ofav$ID<-paste(Ofav$Sample.Name,Ofav$File.Name, sep = ".")
    
  # Check the number of times that a sigle sample was ran  
    ReRunA <- Ofav[duplicated(Ofav$Sample.Name),]
    
    n_RunA <- data.frame(table(Ofav$Sample.Name))
    colnames(n_RunA)<-c("Sample.Name","RanA")
    Ofav<-join(Ofav, n_RunA, type = "left")
    
    DuplicatesA <- Ofav[(Ofav$RanA>1),]
    write.csv(DuplicatesA, file = 'DuplicatesA.csv')
    
```

* Look at the duplicates in 'DuplicatesA.csv' and select the best runs per core. In the DuplicatesA file I delete the good runs and keep the bad ones that are saved a new file called "ToRem1.csv". I use that file to remove them from the general data frame.

* Finally check for duplicates again (you should not have) and summarize the data to check that it makes sense.

```{r}
  # Remove bad replicates
    ToRem2<-read.csv("ToRem2.csv") 
    # 05/04/17
    Ofav<-Ofav[!(Ofav$ID %in% ToRem2$ID),]
    
  # Check for replicates again--should have none
    
    n_RunB <- data.frame(table(Ofav$Sample.Name))
    ReRunB <- Ofav[duplicated(Ofav$Sample.Name),]
    colnames(n_RunB)<-c("Sample.Name","RanB")
    Ofav<-join(Ofav, n_RunB, type = "left")
    
    
  # List of dupplicated samples, should have 0 rows now
    DuplicatesB <- Ofav[(Ofav$RanB>1),]
    #write.csv(DuplicatesB, file = 'DuplicatesB.csv')
```     

5. Check the data!

```{r}
      summary(Ofav)
```

6. Export data if a want local backup

```{r}
    # write.csv(Ofav, file = 'OfavSH.csv')
```
   
    
#---------------------------------------------------------------------------------
## 3. DATA ANALYSIS

    
### 1.Define labeling and factors for each Ofav core sampled: 
    
* Import treatment info from last Lauren's file 

* Merge SH ratios and Treatment information

```{r Treatments}
    Treatments<-read.csv("TGFB-Final_AP.csv", header = TRUE)

    Ofav.data<-join(Ofav, Treatments, by = "Sample.Name", 
                  type = "full", match = "all")
    Ofav.data<-subset ( Ofav.data, select = -c
                        (C.reps, D.reps, Ofav.reps, C.SH, HostCopy, Parent, Original.Tank, X4.AquarumPlacement, Tank, Aquarium) )
   # Export data if a want local backup
    write.csv(Ofav.data, file = 'OfavData.csv')
```

### 2. Exploratory graphs 

#### 2.1 Histograms

Histogram of log10(SH) using all data

```{r Hsstograms}
  HistoSH<-qplot(tot.SH, data=Ofav.data, binwidth=0.02)  
  # HistoSH
    
  HistoL_SH<-qplot(Log_tot.SH_c, data=Ofav.data, binwidth=0.15)  
  HistoL_SH
  
```
    
Histogram of log10(SH) by Combo (Nutrient Treatment * LGTB Treatment)

```{r}
    hist_Tre <- ggplot(Ofav.data, aes(x=Log_tot.SH_c, fill=Treatment))
    
    hist_Tre + 
      geom_density(alpha = 0.2) + facet_grid(Treatment.2~.)
```

#### 2.2 Boxplot: Median,lower and upper hinges=first and third quartiles (the 25th and 75th percentiles) 
+ Mean -+ bootstrap confidence intervals of 95%

SH by Nutrients: 
  
```{r}

    logSHTreatment_C <- ggplot(Ofav.data, aes(factor(Treatment), Log_tot.SH_c)) +
        geom_boxplot(aes(fill=factor(Treatment)), alpha = 0.5) +
        stat_summary(fun.data = "mean_cl_boot",geom = "errorbar", width = 0.2)+
        stat_summary(fun.y=mean, geom="point", size =3, alpha=0.5) +
        geom_jitter(width = 0.2, aes(colour =factor(Colony)))
      logSHTreatment_C
      
     # logSHTreatment <- ggplot(Ofav.data, aes(factor(Treatment), Log_tot.SH)) +
      #  geom_boxplot(aes(fill=factor(Treatment))) +
      # stat_summary(fun.data = "mean_cl_boot",geom = "errorbar", width = 0.2)+
      # stat_summary(fun.y=mean, geom="point", size =3, alpha=0.5) +
      # geom_jitter(width = 0.2, aes(colour =factor(Colony)))
    #  logSHTreatment
```

Boxplot SH by Nutrients * TGFB: 

```{r}
   # logSHTreatment +
     # facet_grid ((~Treatment.2))

 logSHTreatment_C +
      facet_grid ((~Treatment.2))
```
   

* I think that 6 may be have lower values in the NH4 treatment because of the lower time with nutrients after mortallity?

  
Possible colony effect
  
```{r }
   # logSHTreatment +
     # facet_grid (~Colony)
    logSHTreatment_C +
      facet_grid (~Colony)
```

Maybe consider to exclude colony 6 or to treat the SH cell ratio as a continous variable instead as a high/low category. See ANOVA at the end.

```{r}
    #logSHTreatment +
    #  facet_grid (Treatment.2~Colony)

    logSHTreatment_C +
      facet_grid (Treatment.2~Colony)

```
  
Possible problematic combinations: 

* Colony 6 in Control
* Colony 4 in Inhibitor
* Colony 4 in LPS?

Explore later, check weird SH Values and new introduced cores in the NH4 treatmet.


### 3. ANOVAs
    
#### 1. One way ANOVA to test the effect of nutrients alone: 
    
```{r}
    OnwWayANOVA<-aov(Log_tot.SH_c~Treatment, data=Ofav.data)
        summary(OnwWayANOVA)
      
```
Great! Nutrients increased the SH cell ratio!

#### 2. Two Way ANOVA to test effects of Nutrients (Treatment) * TGFB (Treatment.2)

```{r}
    
    TwoWayANOVA<-aov(Log_tot.SH_c ~ Treatment+Treatment.2 , data=Ofav.data)
        summary(TwoWayANOVA)
        drop1(TwoWayANOVA,~.,test="F")
        
```

ATENTION!!!! summary(ANOVA) and drop1(ANOVA) give different p values for the model. I have to investigate what the difference is, but in both cases Treatment (Nutrients) has a significant effect on the SH cell ratio while Treatment.2 (TGFB) does not, nor the interaction. Great!!!

However, I am afraid that the differences in the SH cell ratio in "Inhibitor" are not very strong.


#### 3. One Way ANOVA within factor to test effects of Nutrients (Treatment)/ Colony
  
```{r Models}

  library(lmerTest)
  library(effects)


   OneFactorLMcolony<-lmer(Log_tot.SH_c ~ Treatment + (1|Colony),data=Ofav.data)
        anova(OneFactorLMcolony)
        summary(OneFactorLMcolony)
        
        plot(Effect(c("Treatment"), OneFactorLMcolony), x.var="Treatment", multiline=T, ci.style="bars")
        
```

```{r}
    #ToRunCopy<-anti_join(Ofav.data, OfavCopy, by = "Sample.Name")
    #git pull -u origin master
    #git push -u origin master

```
